\vspace{-2cm} Government websites as data: A methodological pipeline for collection, processing, and text analysis

### 2018-01-30 paper/manuscript.tex
justification for and explanation of the ground truth test

>>>>

\section{Ground truth test}
In the realm of public administration, the notion that the partisan leaning of mayors might have an effect on how they run their cities is still frowned upon to some extent. Perceived more as managers than politicians, they have been portrayed as the last bastion of non-partisanship in America, and in many cases, also style themselves that way \citep{Dovere2018}. However, the aspirations some mayors have shown towards higher offices - in some cases even the presidency - reveal that they are not quite as above the fray as some may believe them to be. One of the most vicious and blatantly partisan cleavages in current U.S. politics - the debate surrounding sanctuary cities - has seen mayors in a central role. Research into local politics has shown that partisan elections consistently have greater turnout \citep{Schaffner2001}. When voters are denied this cue, they make use of other, and considerably more irrational heuristics, such as name, gender, or occupation of the contenders. Consequently it only makes sense for any office-seeking politician to emphasize their partisanship. Finally, decades of research in political psychology have consistently shown that no matter how hard we try, humans are simply incapable of escaping our partisan biases, a finding that is especially pronounced among elites \citep{Hatemi2011}.

In an effort to underline this fact and remove any doubt about the fact that the partisanship of mayors colors their decision-making, we conduct a ground truth test between our main corpus - the websites of cities - and a second, decidedly more partisan set of texts: the campaign websites of these mayors. As noted above, partisanship has been shown to be a powerful driving force even in local politics, and mayors are incentivized to exploit it. Consequently they are very likely to emphasize conservative/liberal values on this platform. If there is a greater correlation in word use between the cities managed by a party and the campaign websites of its mayors than with those of the other party, evidence for the partisanship of city websites can be established.

Using the same methods as described for our main corpus, we have gathered these sites and then concatenated all of the documents belonging to mayors of the two parties into one ground truth document each. We do the same for the city documents, and the compare the four document collections using cosine similarity. This measure is the cosine between the angle of two vectors, in this case the frequencies of all words in the two vocabularies. Compared to a simple euclidean distance, this has the advantage of accounting for the fact that the two corpora being compared are not necessarily of the same length. The cosine measure between two documents ranges between 0 and 1, 0 signifying absolutely no correlation, and 1 perfect overlap. Figure \ref{groundtruth} shows the result of this test. The expectation is for a greater similarity between Republican cities and the Republican ground truth, than Republican cities and Democratic ground truth - and vice versa. At present however, this does not appear to be the case, presumably because the Republican ground truth consists of 8 documents, and the Democratic one of 290.

%ground truth test
\begin{figure}[htp]
	\centering % Using \begin{figure*} makes the figure take up the entire width of the page
	\includegraphics[width=\linewidth]{figures/groundtruth_corrplot.png}
	\caption{Ground truth test. The values are cosine similarities between a pair of document collections.}
	\label{groundtruth}
\end{figure}
